|

### 2021-09-29 05-Chp-IPD-Analysis.Rmd
Back to working version General clean-up Proper counting of words included in tables
End of day:

---
#########################################
# options for knitting a single chapter #
#########################################
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: default
  bookdown::html_document2: default
documentclass: book
---

@@

@@

As part

As part of the IPD analysis, relevant cohort studies identified through the systematic review detailed in Chapters \@ref(sys-rev-heading) were. In the first instance, the corresponding author on the main man

@@

In addition, to assess the results for  the Dementia Platform UK common-access procedure was utilised. In short, this approach is intended to make it easier to access data from exisiting dementia cohorts through a centralised application process.

Talk about how different covariates were managed across cohorts,

Lipids reported/available as a continuous measure, and were not cross-sectional.

Due to the limited time-frame, studies making use of population-level electronic health records, which often require an entire project proposal, were ineligible due to the time and cost involved in applying. The one exception to this was data from the CPRD, which we already had access to via the study reported in Chapter \'ref(cprd-heading).

Talk about missing data were handled <!-- IDEA See Sean's thesis for some additional detail on this -->

@@

Risk of bias assessment was performed for each of the included cohorts using the relevant tool described in Section \@ref(risk-of-bias).

There is some concerns about performing risk of bias assessments on your own analysis, and so

@@

@@

<!-- Be clear on the different types of databases, e.g. those where users are followed up regularly vs those where participants are captured via EHR -->

@@

All analysis were standardised by changes in 1-SD of the exposure variable. This was done to aid interpretability of the outcome.

Similary,

Clustering within studies was accounted for, given the evidence that ...[@abo-zaid2013]

Stratified by age-at-entry and sex, and ethnicity where possible.

Hazard ratios were utilised to

In order to investigate the interaction of sex with

A discrete proportional hazards model was employed to account for the interval censoring introduced by design o fthe longitudinal cohort studies.[@wang2017]

@@

Where results were comparably to previously published estimates, these were compared and reasons for any discrepancies discussed.

The analysis were carried out in R (Version...) using the following packages: `survival`[@survival-book], `metafor`[@viechtbauer2010a]

@@

Where any included data source had previously been analysed and results for the association between lipid levels and dementia reported, these were compared to the results of this analysis.

@@

For the streamlined DPUK process, where a dedicated project manager liases with cohort owners, the overall response rate was higher, result in access to three cohorts. However, even using th

@@

I did not request data from several DPUK cohorts due to the information on the online system indicating that the study was relatively new (and so was yet to collect >1 wave of data, as in the case of the XXXX cohort) or the online data dictionary indicating that the exposure variables of interest were not recorded.

@@

The three datasets included are described in detail in the following section, and are summarised in Table \@ref(tab:cohortOverview-table).

@@

<!-- TODO Update table with SD for age, and check why number of events in EPIC is less than expected (5 vs 8)-->

@@

Cholesterol measures (total, LDL-c, HDL-c and triglycerides) were measured at baseline in 1979-1983. 
As the study population has aged, additional outomes. Of particular relevance to this analysis, from Phase III (1989-1993) onwards, a battery of congnitive tests were introduced.

@@

Additionally, ther number of participants is larger in this analysis as there is no restriction on the level which lipids should be in order to be included in the analysis.

@@

Has only 8 events

```{r tests, }

```

@@

This data source was analysed in one of the included studies identified by the systematic review presented in Chapter \@ref(sys-rev-heading).[@tynkkynen2018]

This gave

@@

### Cohorts provided data but ultimately excluded

@@

Likley had dementia at baseline

@@

### Cohorts identified the systematic review but excluded from the IPD analysis

### Cohorts approach but received no answer

@@

@@

The advantage of this approach is that it reduces the liklihood of errors in model mis-speciication if needing to change variables names from cohort to cohort.

@@

@@

@@

### Resources

@@

* Problems with getting data from studies
*

Sharing data has several legal, ethical and logistical challenges, and can often

This is in theory what the DPUK was built to address. However, despite claims to a streamlined process, the response rate amond cohorts participating

Describe your experience of trying to access the DPUK - while a great resources, frustrating at times. The portal would benefit from a centralised management system and

Even with the guaranteed data access afforded by the DPUK, accessing sufficient data were

Of the XXX cohorts requested from the DPUK, allowing for a year lag since application, a response was received from only XXX (XXX%), while of those responding a minority

@@

>>>>

---
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: default
  bookdown::html_document2: default
documentclass: book
---

@@

Eligible data sources were defined as those that contained lipid measurements and  ipids reported/available as a continuous measure, Data sources which were cross-sectional, either by design or due to the available data (e.g. a study conducted multiple waves, but onyl data from a single wave could be accessed) were excluded. Similarly, due to the limited time-frame, studies making use of population-level electronic health records, which often require an extensive project proposals in order to gain access to the data, were ineligible due to the time and cost involved in applying. The one exception to this was data from the CPRD, which I already had access to as part of the analysis reported in Chapter \ref(cprd-heading).

@@

As part of the IPD analysis, eligible observational prospective cohort studies identified through the systematic review detailed in Chapters \@ref(sys-rev-heading) were approached for data access.

In the first instance, the first author on the main study report was emailed (see Appendix \@ref() for a copy of the text and documentation sent). If this did not elict a response, the last/corresponding author was contacted, on the basis that the first author may have been a more junior member.

examining the effect of lipid levels

@@

In addition, in an attempt to supplement the analysis with

In addition, to assess the results for  the Dementia Platform UK common-access procedure was utilised. In short, this approach is intended to make it easier to access data from existing dementia cohorts through a centralised application process.

@@

Risk of bias assessment was performed for each of the included cohorts using the risk-of-bias assessment tool for non-randomised studies of exposures introduced previously (for a more detailed discussion of this tool, see Section \@ref(rob-tools-nrse)).

@@

<!-- TODO This was taken from Chapter 4, so may need to rephrase slightly -->
<!-- IDEA See Sean's thesis for some additional detail on this -->

In this analysis, missing data were handled using a multiple imputation approach. Variables with missing observations were identified, and 20 imputed datasets were created.[@sterne2009] Nominal variables with missing values were modelled using multinomial logistic regression, while continuous variables were modelled using linear regression. As per best practice, all variables used in the analytic model, including the outcome, were included in the imputation model.[@moons2006using] Imputation was performed using the MICE (Multiple Imputation by Chained Equations) using the XXX command in R.

Using multiple imputation to handle missing data is an alternative to a "complete case" approach,[@pigott2001review] where participants missing any covariate are dropped from the dataset. As a sensitivity analysis, I preformed and compared the results of both methods,[@hughes2019] to investigate the impact of multiple imputation on the results.

<!-- TODO Need to talk about how missing variables were handled across studies -->

@@

_Individual patient data approached__

Two stage versus one-stage approach

__Statistical methods__

Under a two-stage approach, estimates for the effect of each lipid fraction on available dementia outcomes in each study were first calculated for each data source. A common model, adjusted for identical covariates, were obtained using a common model.

To aid interpretability of the outcome, all analyses were standardised by changes in 1-SD of the exposure variable. Clustering within studies was accounted for, given the evidence that ...[@abo-zaid2013]

Hazard ratios were used to quantify the effect of a 1-SD lower exposure to a lipid fraction on dementia outcomes. A discrete proportional hazards model was employed to account for the interval censoring introduced by design of the longitudinal cohort studies.[@wang2017]

The estimates from each data source were then combined using a random-effects meta-analysis

@@

__Investigating effect of patient-level covariates__

@@

@@

Where an included data source had previously been analysed and results for the association between lipid levels and dementia reported, these were compared to the results of this analysis. If discrepancies were identified, these were investigate.

__Combining published aggregate data and individual participant data__

Where a data source did not provided their individual participant data, but stratified by a patient characteristic of interest in the published data, the agregate date and individual data were combined.

<!-- TODO Need more on this method here -->

__Author statements on data availability__

As a sensitivity analysis, I examined the data sharing statements for each using a predefined categories used in a similar previous analysis.[@mcguinness2020DAScomparison]

<!-- TODO This should have a plot showing how many over time had good vs bad DAS -->

<!-- TODO In addition, need plot similar to study design over time in Chapter 3, showing cohort responses over time -->

@@

For the streamlined DPUK process, where a dedicated project manager liaises with cohort owners on the applicants behalf, the overall response rate was higher, resulting in access to three cohorts. However, even using this dedicated access approach, a response was obtained for less than half of the approach cohorts

@@

@@

The four data sources used in this analysis are summarised in Table \@ref(tab:cohortOverview-table) and are described in detail in the following sections.

@@

<!-- TODO Update table with SD for age, and types of outcomes, and follow-up, and check why number of events in EPIC is less than expected (5 vs 8)-->

@@

Cholesterol measures (total, LDL-c, HDL-c and triglyceride) were measured at baseline in 1979-1983. 
As the study population has aged, additional outcomes. Of particular relevance to this analysis, from Phase III (1989-1993) onwards, a battery of cognitive tests were introduced.

@@

Additionally, other number of participants is larger in this analysis as there is no restriction on the level which lipids should be in order to be included in the analysis.

@@

The added evidental value of the EPIC cohort is small, given the relatively small number of participants and the fact that the cohort has only 8 dementia events.

```{r tests}
```

@@

This data source was analysed in one of the included studies identified by the systematic review presented in Chapter \@ref(sys-rev-heading),[@tynkkynen2018] meaning that a comparison between the published result and the analysis reported here was was possible.

@@

### Excluded data sources

#### Cohorts identified the systematic review but excluded from the IPD analysis

Lothian Birth Cohort - 1936

Taiwan, etc.

#### Cohorts identified in the DPUK but excluded from the IPD analysis

Several additional cohorts are present in the DPUK collection but were not approach with a request for data acccess. This was primarily due to the information on the online system indicating that the study was relatively new (and so was yet to collect >1 wave of data, as in the case of the XXXX cohort) or the online data dictionary indicating that the exposure variables of interest were not recorded.

#### Cohorts providing data but ultimately excluded

@@

Likley had dementia at baseline. However, on further investigation of the cohort, it was discovered that

@@

#### Cohorts approach but received no answer

@@

These included studies with  a dedicated access process or data access committee. A cohort of particular interest, having been used in multiple studies identified by the systematic review is the Three City Study in France. This study was approach at multiple points over the curse of this analysis, and no answer was received.

This was also a problem for cohorts approach via the DPUK - half of the cohorts applied to did not respond to the application within a year.

@@

The advantage of this approach is that it reduces the likelihood of errors in model mis-specification if needing to change variables names from cohort to cohort.

@@

# Results

<!----------------------------------------------------------------------------->
(ref:tcDementiaMain-cap) __Random effects meta-analysis of estimates from included data sources:__

(ref:tcDementiaMain-scap) IPD analysis of total cholesterol and dementia

```{r tcDementiaMain, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:tcDementiaMain-cap)', out.width='100%', fig.scap='(ref:tcDementiaMain-scap)'}
knitr::include_graphics(file.path("figures/ipd/TC.png"))
```
<!----------------------------------------------------------------------------->

@@

#### Risk of bias assessment

There is some concerns about performing risk of bias assessments on your own analysis, and so two secondary reviewers

#### Opportunistic inclusion of CPRD data

@@

### Strengths

While this analysis did not manage to systematically obtain and analyse a large proporiton of identified data, it did enable an analysis of two previously unanalysed datasets - the CaPS study and the - providing additional data points in

it also attempted to make use

@@

* Problems with getting data from studies
* Problems with current methods of tracking people down via . Particularly when studies were older, several cases of the address/contact details for the corresponding author being out of date. Matches with decrease in availability as the studies get older.

__Many authors claim that their data can be made “available on request”, despite previous work establishing that these statements are demonstrably untrue in the majority of cases—that when data is requested, it is not actually made available. [@krawczyk2012; @miyakawa2020; @naudet2018] Additionally, previous work found that the availability of data “available on request” declines with article age, indicating that this approach is not a valid long term option for data sharing.[@vines2014]__ <!-- TODO taken from DAS paper, so will need to change language -->

Qualitatively, investigation of a subset of cohorts on an _ad-hoc_ basis reveal that several corresponding/first authors were no longer at the same institution as when the study was reported, and as a result, were unlikely to have access to the institutional email address listed on the study publication. All of these are arguments towards better reporting of data availability, ownership and

Even with the help of the streamlined application process afforded by the DPUK, accessing sufficient data was a challenge. Sharing data has several legal, ethical and logistical challenges, and can often taken several years to obtain. This is in theory what the DPUK was built to address. However, despite claims to a streamlined process, the response rate among cohorts a year after application was 50%. In addition, some cohorts responded that that the proposed study question was already under investigation by another group, and that they would not share the data on this basis. In light of this, a centralised database of ongoing analysis being performed in DPUK would be of enormous help. Finally, the DPUK should more clearly distinguish between those cohorts that are "DPUK native" versus externally hosted, as the response time for externally hosted cohorts is likely to be much longer.

@@

#### Aims/Resources

### 2021-09-29 _main.Rmd
End of day:

#### Non-randomised studies of interventions/exposures

@@

@@

@@

Has only 8 events

```{r tests, }

@@

<!-- ### Triangulation across evidence sources -->

@@

# Triangulation {#tri-heading}

@@

>>>>

#### Non-randomised studies of interventions/exposures {#rob-tools-nrse}

@@

---
output:
  bookdown::pdf_document2:
    template: templates/brief_template.tex
  bookdown::word_document2: default
  bookdown::html_document2: default
documentclass: book
---

@@

<!-- TODO This was taken from Chapter 4, so may need to rephrase slightly -->
<!-- IDEA See Sean's thesis for some additional detail on this -->

In this analysis, missing data were handled using a multiple imputation approach. Variables with missing observations were identified, and 20 imputed datasets were created.[@sterne2009] Nominal variables with missing values were modelled using multinomial logistic regression, while continuous variables were modelled using linear regression. As per best practice, all variables used in the analytic model, including the outcome, were included in the imputation model.[@moons2006using] Imputation was performed using the MICE (Multiple Imputation by Chained Equations) using the XXX command in R.

<!-- TODO Need to talk about how missing variables were handled across studies -->

@@

The added evidental value of the EPIC cohort is small, given the relatively small number of participants and the fact that the cohort has only 8 dementia events.

```{r tests}
```

@@

<!-- ### Triangulation  -->

@@

# Aetiological triangulation across evidence sources {#tri-heading}

@@

<!-- TODO These citations are not rendering properly - I think because bookdown using this latex format only allows 1 .bib file. -->

### 2021-09-29 front-and-back-matter/_00-introduction.Rmd
End of day:

```{r word-count, echo=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)

options(scipen=999)

@@

```{r word-count-fig, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(mapping = aes(x=date,y=words)) +
  geom_line(data = target_df, colour = "black", linetype = 2) +
  geom_hline(yintercept = 40000, colour = "black", linetype = 2) +
  geom_vline(xintercept = as.numeric(as.Date("20211116", format="%Y%m%d")), colour = "black", linetype = 2) +
  geom_vline(xintercept = as.numeric(as.Date("20211216", format="%Y%m%d")), colour = "red") +
  scale_x_date(limits = c(as.Date("20201124", format="%Y%m%d"),as.Date("20211231", format="%Y%m%d"))) +
  geom_step(data = words, colour = "red")

>>>>

```{r word-count, echo=FALSE, warning=FALSE, message=FALSE}
options(scipen=999)

@@

```{r word-count-fig, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(mapping = aes(x=date,y=words)) +
  geom_line(data = target_df, colour = "black", linetype = 2) +
  geom_hline(yintercept = 40000, colour = "black", linetype = 2) +
  geom_vline(xintercept = as.numeric(as.Date("20211116", format="%Y%m%d")), colour = "black", linetype = 2) +
  geom_vline(xintercept = as.numeric(as.Date("20211216", format="%Y%m%d")), colour = "red") +
  scale_x_date(limits = c(as.Date("20201124", format="%Y%m%d"),as.Date("20211231", format="%Y%m%d"))) +
  geom_step(data = words, colour = "red") +
  geom_text(aes(x = as.Date("20210910", format="%Y%m%d"), y = 30000, label = "Tables"))
